\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities

### 2020-12-27 popl-2019-journal-v/discussion.tex
Fix another reviewer remark.

\begin{jversion}
  A compiler is secure when it enforces the properties of high-level languages which begs the question what properties we should enforce.
  When it comes to fully-abstract compilation, then the answer is that all properties in the
  high-level language should be enforced (or at least the equivalences they imply), so the real question is what high-level language we would like.
  \stktokens{} ensures a standard call-return control-flow, but if we want a different kind of control-flow, for instance call/cc or C's \texttt{longjmp}, then we need to come up with a different enforcement scheme.
  Further, many high-level languages have exceptions as another form of control-flow not (yet) supported by \stktokens{}.
  This goes to show how we must consider what high-level language we want in order to answer the question of what properties we must enforce to get a fully-abstract compiler.
  % Lau: Say something about address hiding?

>>>>

\begin{jversion}
  A compiler is secure when it enforces the properties of high-level languages which begs the question what properties we should enforce.
  When it comes to fully-abstract compilation, then the answer is that all properties in the
  high-level language should be enforced (or at least the equivalences they imply\footnote{See Section~\ref{sec:related-work} for a discussion of secure compilation properties that require preservation of more general language properties than equivalences.}) must be preserved in the target language, either by translating them to similar abstractions in the target language or by using the available target language features to enforce the source language abstraction.
  In our case, \srccm{} semantics offers a native stack with LSE and WBCF, but this abstraction does not natively exist in \trgcm{}.
  In order to enforce the abstraction on \trgcm{}, \stktokens{} can be used.
  Theorem~\ref{thm:full-abstraction} proves that \stktokens{} enforces these abstraction properties in a way that behaviorally matches \srccm{} which means that it enforces LSE and WBCF.
), so the real question is what high-level language we would like.
  \stktokens{} ensures a standard call-return control-flow, but if we want a different kind of control-flow, for instance call/cc or C's \texttt{longjmp}, then we need to come up with a different enforcement scheme.
  Further, many high-level languages have exceptions as another form of control-flow not (yet) supported by \stktokens{}.
  This goes to show how we must consider what high-level language we want in order to answer the question of what properties we must enforce to get a fully-abstract compiler.
  % Lau: Say something about address hiding?

### 2020-12-27 popl-2019-journal-v/paper.tex
Fix another reviewer remark.

% Programs
We can now define the notion of a program as well as a context.
\begin{definition}[Programs and Contexts]
  \label{def:program-and-context}
  A \emph{program} is a component $(\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$ with an empty import list.
  A \emph{context} for a component $\comp$ is a component $\comp'$ such that $\comp \bowtie \comp'$ is a program.
\end{definition}
% Getting an executable from a component - postpone this to later, but mention:
% write xor exec
How a program is initialised to create an executable configuration is discussed in Section~\ref{sec:form-secur-with}.
% We will, however, at this point mention that all programs will be instrumented as Write-XOR-Execute,{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
%Further, we will only consider a specific class of well-formed components, but we will get back to this.
Some simplifications have been made in this presentation of \trgcm{}. See \citet{technical_report_popl} for details.

@@

Contextual equivalence roughly says that two components behave the same no matter what context we plug them into.
%The remaining defintions
\begin{definition}[Plugging a component into a context]
  \label{def:comp-context-plugging}
  When $\var{comp'}$ is a context for component $\var{comp}$ and $\var{comp}' \bowtie \var{comp} \rightsquigarrow \Phi$, 
  then we write $\plug{\var{comp'}}{\var{comp}}$ for the execution configuration $\Phi$.
\end{definition}
\begin{definition}[\trgcm{} and \srccm{} contextual equivalence]
  \label{def:contextual-equivalence}
  ~
  \begin{description}
  \item[On \srccm{}], we define that $\src{\comp_1 \sconeq \comp_2}$ iff
  \begin{equation*}
    \forall \src{\context} \ldotp  \wdjud[\emptyset]{\src{\context}} \Rightarrow \src{\plug{\context}{\comp_1} \sterm{\ta[,1],\stkb_1}} \Leftrightarrow \src{\plug{\context}{\comp_2} \sterm{\ta[,2],\stkb_2}}
  \end{equation*}
  with $\src{\ta[,i]} = \src{\dom(\comp_i.\mscode)}$.
  \item[On \trgcm{}], we define that $\trg{\comp_1 \tconeq \comp_2}$ iff
  \begin{equation*}
    \forall \trg{\context} \ldotp \wdjud[\emptyset]{\trg{\context}} \Rightarrow \trg{\plug{\context}{\comp_1} \term} \Leftrightarrow \trg{\plug{\context}{\comp_2} \term}
  \end{equation*}
\end{description}
where $\Phi\sterm[i]{\src{\ta,\stkb}} \text{ iff } \Phi \nstep[i]{\src{\ta,\stkb}} \halted \quad$ and $\quad \Phi\sterm{\src{\ta,\stkb}} \defeq \exists i \ldotp \sterm[i]{\src{\ta,\stkb}}$
\end{definition}
With the above defined, we are almost ready to state our full-abstraction, and all that remains is the compiler we claim to be fully-abstract.
We only care about the well-formed components, and they sport none of the new syntactic constructs \srccm{} adds to \trgcm{}.
This means that the compilation from \srccm{} components to \trgcm{} components is simply the identity function.
% Full abstraction thm
\begin{theorem}
  \label{thm:full-abstraction}
  For reasonable, well-formed components $\comp_1$ and $\comp_2$, we have
  \begin{gather*}
    \src{\comp_1} \sconeq \src{\comp_2} \quad \Leftrightarrow \quad    \src{\comp_1} \tconeq \src{\comp_2} \qedhere
  \end{gather*}
\end{theorem}
Readers unfamiliar with fully-abstract compilation may wonder why Theorem~\ref{thm:full-abstraction} proves that \stktokens{} guarantees LSE and WBCF.
Generally speaking, behavioral equivalences are preserved and reflected by fully-abstract compilers.
This means that any property the source language has must somehow be there after compilation, whether or not it is a property of the target language.
If the source language has a property that the target language doesn't have, then a compiled source program must use the available target language features to emulate the source language property in a way that exactly matches behaviorally.
In our case, LSE and WBCF are built into the semantics of \srccm{}, but they are not properties of \trgcm{}.
In order to enforce these properties, components on \trgcm{} use \stktokens{}.
Theorem~\ref{thm:full-abstraction} proves that \stktokens{} enforces these properties in a way that behaviorally matches \srccm{} which means that it enforces LSE and WBCF.

>>>>

% Programs
We can now define the notion of a program as well as a context.
\begin{definition}[Programs and Contexts]
  \label{def:program-and-context}
  A \emph{program} is a component $(\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$ with an empty import list.
  A \emph{context} for a component $\comp$ is a component $\comp'$ such that $\comp \bowtie \comp'$ is a program.
\end{definition}
% Getting an executable from a component - postpone this to later, but mention:
% write xor exec
How a program is initialised to create an executable configuration is discussed in Section~\ref{sec:form-secur-with}.
% We will, however, at this point mention that all programs will be instrumented as Write-XOR-Execute,{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
%Further, we will only consider a specific class of well-formed components, but we will get back to this.
Some simplifications have been made in this presentation of \trgcm{}.
These simplifications concern minor details, for example, we've omitted the fact that \texttt{move}s to the \texttt{pc} register are disallowed.
Full details can be checked in \citet{technical_report_popl}.

@@

Contextual equivalence roughly says that two components behave the same no matter what context we plug them into.
%The remaining defintions
\begin{definition}[Plugging a component into a context]
  \label{def:comp-context-plugging}
  When $\var{comp'}$ is a context for component $\var{comp}$ and $\var{comp}' \bowtie \var{comp} \rightsquigarrow \Phi$, 
  then we write $\plug{\var{comp'}}{\var{comp}}$ for the execution configuration $\Phi$.
\end{definition}
\begin{definition}[\trgcm{} and \srccm{} contextual equivalence]
  \label{def:contextual-equivalence}
  ~
  \begin{description}
  \item[On \srccm{}], we define that $\src{\comp_1 \sconeq \comp_2}$ iff
  \begin{equation*}
    \forall \src{\context} \ldotp  \wdjud[\emptyset]{\src{\context}} \Rightarrow \src{\plug{\context}{\comp_1} \sterm{\ta[,1],\stkb_1}} \Leftrightarrow \src{\plug{\context}{\comp_2} \sterm{\ta[,2],\stkb_2}}
  \end{equation*}
  with $\src{\ta[,i]} = \src{\dom(\comp_i.\mscode)}$.
  \item[On \trgcm{}], we define that $\trg{\comp_1 \tconeq \comp_2}$ iff
  \begin{equation*}
    \forall \trg{\context} \ldotp \wdjud[\emptyset]{\trg{\context}} \Rightarrow \trg{\plug{\context}{\comp_1} \term} \Leftrightarrow \trg{\plug{\context}{\comp_2} \term}
  \end{equation*}
\end{description}
where $\Phi\sterm[i]{\src{\ta,\stkb}} \text{ iff } \Phi \nstep[i]{\src{\ta,\stkb}} \halted \quad$ and $\quad \Phi\sterm{\src{\ta,\stkb}} \defeq \exists i \ldotp \sterm[i]{\src{\ta,\stkb}}$
\end{definition}
With the above defined, we are almost ready to state our full-abstraction, and all that remains is the compiler we claim to be fully-abstract.
We only care about the well-formed components, and they sport none of the new syntactic constructs \srccm{} adds to \trgcm{}.
This means that the compilation from \srccm{} components to \trgcm{} components is simply the identity function.
% Full abstraction thm
\begin{theorem}
  \label{thm:full-abstraction}
  For reasonable, well-formed components $\comp_1$ and $\comp_2$, we have
  \begin{gather*}
    \src{\comp_1} \sconeq \src{\comp_2} \quad \Leftrightarrow \quad    \src{\comp_1} \tconeq \src{\comp_2} \qedhere
  \end{gather*}
\end{theorem}
Readers unfamiliar with fully-abstract compilation may wonder why Theorem~\ref{thm:full-abstraction} proves that \stktokens{} guarantees LSE and WBCF.
Generally speaking, behavioral equivalences are preserved and reflected by fully-abstract compilers.
This means that source language abstractions (or at least the equivalences they imply\footnote{See Section~\ref{sec:related-work} for a discussion of secure compilation properties that require preservation of more general language properties than equivalences.}) must be preserved in the target language, either by translating them to similar abstractions in the target language or by using the available target language features to enforce the source language abstraction.
In our case, \srccm{} semantics offers a native stack with LSE and WBCF, but this abstraction does not natively exist in \trgcm{}.
In order to enforce the abstraction on \trgcm{}, \stktokens{} can be used.
Theorem~\ref{thm:full-abstraction} proves that \stktokens{} enforces these abstraction properties in a way that behaviorally matches \srccm{} which means that it enforces LSE and WBCF.
